Inference of Gravitational Lens Time Delays Using Gaussian Processes

### 2012-05-30 doc/ms.tex
Random edits

\section{The Model}

@@

The probability distribution for the observed data $\mathbf{y}$ given the
model parameters $\boldsymbol{\theta}$ is a multivariate Gaussian:
\begin{equation}
p(\mathbf{y} | \theta) = \frac{1}{\sqrt{(2\pi)^n \det \mathbf{C}}} \exp\left[-\frac{1}{2}(\mathbf{y} - \mathbf{m})^T\mathbf{C}^{-1}(\mathbf{y} - \mathbf{m})\right]
\end{equation}
where $\mathbf{m}$ is the vector of mean values, and $\mathbf{C}$ is the
covariance matrix of the data, formed by evaluating the covariance function of
the GP at the times of the observations, and then adding diagonal components
to describe the observational errors. That is to say,

\begin{equation}
\mathbf{C} = \textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right] + \mathbf{C}_{\rm noise}.
\end{equation}
{\bf TODO: WRITE THIS PROPERLY}

@@

A common situation in astronomy is that final data products, such as light
curves, are returned, with error bars, from a pipeline. However, such error
bars are not always trustworthy, as they may have been underestimated due to
modelling assumptions and approximations in the pipeline. In the context of
light curve modelling, underestimated error bars can be detected if the
scatter in the light curve is better modelled by independent Gaussian noise
than it is by a time-correlated microlensing term. Thus, to improve the
robustness of our method, we include a ``noise boost'' parameter $\kappa$
describing the factor by which the reported error bars should be multiplied.
This is another testable mechanism for accounting for outliers.

In summary, the parameter vector is
\begin{equation}
\left\{ \{\tau\}, \{m\}, \sigma_{\rm QSO}, L_{\rm QSO}, \{\sigma_{\rm micro}\}, \{L_{\rm micro}\}, \alpha_{\rm micro}, \kappa \right\}
\end{equation}
where parameters in braces denote a set (one for each QSO image).  For a
double image lens, the total number of parameters in the model is~12. For a
quad, this number is~20. {\bf CHECK.}  A set of uninformative prior
distributions for these parameters are listed in Table~\ref{priors}; we assign
these priors in the analysis described in the first part of this paper. Later,
we will show how a more informative set of priors can be derived, and used to
improve the precision of the time delay inferences {\bf (WE HOPE)}.

>>>>

\section{The Model for the Inference}

@@

The probability distribution for the observed data $\mathbf{y}$ given the
model parameters $\boldsymbol{\theta}$ is a multivariate Gaussian:
\begin{equation}
p(\mathbf{y} | \theta) = \frac{1}{\sqrt{(2\pi)^n \det \mathbf{C}}}
\exp\left[-\frac{1}{2}
(\mathbf{y} - \mathbf{m})^T\mathbf{C}^{-1}(\mathbf{y} - \mathbf{m})\right]
\end{equation}
where $\mathbf{m}$ is the vector of mean values, and $\mathbf{C}$ is the
covariance matrix of the data, formed by evaluating the covariance function of
the GP at the times of the observations, and then adding diagonal components
to describe the observational errors. That is,
\begin{equation}
\mathbf{C} = \textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right] + \mathbf{C}_{\rm noise}.
\end{equation}
{\bf TODO: WRITE THIS PROPERLY}

@@

A common situation in astronomy is that final data products, such as light
curves, are returned, with error bars, from a pipeline. However, such error
bars are not always trustworthy, as they may have been underestimated due to
modelling assumptions and approximations in the pipeline. In the context of
light curve modelling, underestimated error bars can be detected if the
scatter in the light curve is better modelled by independent Gaussian noise
than it is by a time-correlated microlensing term. Thus, to improve the
robustness of our method, we include a ``noise boost'' parameter $\kappa$
describing the factor by which the reported error bars should be multiplied.
This is another mechanism for accounting for outliers.

In summary, the parameter vector is
\begin{equation}
\left\{ \{\tau\}, \{m\}, \sigma_{\rm QSO}, L_{\rm QSO}, \{\sigma_{\rm micro}\}
\{L_{\rm micro}\}, \alpha_{\rm micro}, \kappa \right\}
\end{equation}
where parameters in braces denote a set (one for each QSO image).  For a
double image lens, the total number of parameters in the model is~12. For a
quad, this number is~20. A set of uninformative prior
distributions for these parameters are listed in Table~\ref{priors}; we assign
these priors in the analysis described in the first part of this paper. Later,
we will show how a more informative set of priors can be derived, and used to
improve the precision of the time delay inferences {\bf (WE HOPE)}.
